### 2019-07-09 TFM.tex
Pictures
Terminamos

\usepackage[left=3.7cm,right=3.7cm,top=3cm,bottom=3cm]{geometry}
\usepackage{graphicx}
\usepackage{amssymb,mathtools, amsmath, amsfonts, amsthm}
%\usepackage{color}
\usepackage{float}
\usepackage{hyperref}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{chngcntr}
\usepackage{cleveref}
\usepackage{pdfpages}

@@

The original goal was to give generalizations for the case of uniform hypergraphs
but the techniques used for that case also allowed for generalizations to 
hypergraphs with multiple edge sets of different sizes.

@@

This way, asymptotically, trees of any kind occur arbitrarily often in a typical
random graph. Given the locality of first order logic up to a fixed 
quantifier rank this is tells us that the only thing that the only way 
to distinguish large random graphs in $G(n,\beta/n)$ 
(up to a fixed quantifier rank) 
is by looking at the neighborhoods of their small cycles. \par

It is given a finite and exhaustive classification of the uni-cyclic
graphs with less than a given radius. After that it is shown that
if two graphs $G_1$ and $G_2$ contain the same number 
of uni-cycles of each type or they
contain a large enough number of those then they are a.a.s equivalent
under first order logic up to some quantifier rank $k$. This is done by giving
a winning strategy for duplicator in the E.F. game of $k$ round played on 
$G_1$ and $G_2$.\par

Finally it is shown that the numbers of uni-cycles of each type contained in
a random graph are asymptotically distributed like independent Poisson variables 
whose means are given by some other nested ``Poisson expressions" that correspond 
to the trees hanging from each uni-cycle.

@@

Given a natural number $n$, we will use the notation
$[n]:= \{1,\dots, n\}$. We will denote by $S_n$
the symmetric group on $[n]$, and by $\Delta_n$ the 
diagonal set $\{(a,a)\in [n]^2 \}$. \par
Given a set $X$, then $S_n$ acts on $X^n$ in an evident way. That is, 
given $g\in S_n$ and $(x_1, \dots, x_n)$ one can define
\[ g \cdot (x_1,\dots,x_n) =(y_1,\dots,y_n), \]
where $y_{g(i)}=x_i$ for all $1\leq i\leq n$.\par
Given $\Phi$ a subgroup of $S_n$ we will denote by $X^n/\Phi$
the orbit set associated to the action of $\Phi$ over $X^n$.\par
We will use the notation $[x_1,\dots,x_n]$ to refer to the 
equivalence class of the $n$-tuple $(x_1,\dots, x_n)$ in 
any sort of quotient $X^n/\Phi$. That is, while the notation
$(x_1,\dots, x_n)$ will be reserved to ordered $n$-tuples, 
$[x_1,\dots,x_n]$ will denote an ordered $n$-tuple modulo the
action of some arbitrary group of permutations. Which group is this 
will depend solely on the ambient set where $[x_1,\dots,x_n]$ is
considered.
\par

\begin{definition}
	Let $n,a\in \N$, let $\Phi$ be a subgroup of $S_a$, and let
	$A$ be a subset 
	\[ A\subseteq [a]^2 \setminus \delta.\]
	The total edge set $\mathcal{H}_{(a,\Phi, A)}(n)$ of size $a$, 
	symmetry group $\Phi$ and
	restrictions $R$ on $n$ elements is the set:
	\[  \mathcal{H}_{(a,\Phi, R)}(n)= ([n]^a/\Phi) \, \,
	\setminus \{\,  [x_1, \dots,x_a] \in [n]^a/\Phi  \, \, 
	| \, \, x_i=x_j \, \text{for some } (i,j)\in R \} \]
\end{definition}

@@

\begin{definition} 
	Let $p=(p_1,\dots, p_c)$, where all $p_i$'s are real numbers
	between $0$ and $1$. 
	The random model $HG(n,p)$ with edge colors $1,\dots,c$,
	sizes $a_1,\dots,a_c$, 
	symmetry groups $\Phi_1,dots,\Phi_c$ 
	and restrictions $A_1,\dots,A_c$, is the one that
	assigns to each graph $G=([n], H_1,\dots, H_c)$
	probability
	\[ \mathrm{Pr}(G)=\prod_{i=1}^{c} p_i^{|H_i|}(1-p_i)^{|\mathcal{H}_{(a_i,\Phi_i,A_i)}(n)|-|H_i|}.	
	\]
	Equivalently, this is the probability space obtained by 
	assigning to each colored edge 
	$e\in \mathcal{H}_{(a_i,\Phi_i,A_i)}(n)$ probability $p_i$ independently. 
\end{definition}

For the rest of the work we will consider 
\begin{itemize}
	\item the total number of colors $c$,	
	\item the sizes $a_1,\dots,a_c$,
	\item the symmetry group $\Phi_1,\dots, \Phi_c$ and,
	\item the restrictions $A_1,\dots,A_l$
\end{itemize}
fixed. When we say ``graph'' from now on what
we will mean is ``hiper-graph with edge colors 
$1,\dots, c$, sizes $a_1,\dots,a_c$, 
symmetry groups $\Phi_1,\dots,\Phi_c$ and 
restrictions $R_1,\dots,R_c$''. \par

Given a graph $G=([n], H_1,\dots, H_c)$ we will denote by
$H_i(G)$ the edge set $H_i$, and by $V[G]$ the vertex set
$[n]$. Also, we will write $H(G)$ to denote the 
disjoint union of colored sets $\cup_{i=1}^c H_i$.
This way, an edge $e\in H(G)$ with color $i$ is an element
$[x_1,\dots,x_{a_i}]\in H_i(G)$, and the $x_i$'s are vertices
belonging to $V(G)$. \par
Given a set of vertices, $X\subseteq V(G)$, we will denote
the by $G[X]$ the induced sub-graph on $X$. \par
As usual, we will sometimes treat edges as sets of vertices
rather than ``tuples modulo the action of some permutation group''. 
This way, expressions like $e_1\cup e_2$ for $e_1, e_2\in H(G)$
will make sense and mean 
``the set of vertices that occupy some place in $e_1$ 
and in $e_2$".\par
Some other times we will treat edges $e\in H(G)$
as sub-graphs of $G$ in the evident way. That is, the subgraph
denoted by $e$ is the one whose vertex set is $e$-i.e., the vertices in $e$-
and whose only edge is $e$. 
This way, when we have some edges $e_1,\dots, e_l\in H(G)$ 
it will make sense to talk
about the subgraph $\cup_{i=1}^l e_i$, which is the graph
whose vertex set is the set of vertices belonging to the $e_i$'s, 
and whose edges are exactly the $e_i$'s. In spite of these
abuses of notation the ``type'' of any ``term'' involving edges 
should be derivable from the context. \par
Another usual abuse of notation we will make is to sometimes
treat graphs as their underlying vertex sets. Hence,
expressions defined for sets of vertices will also be defined 
for graphs.

@@

Graphs are $\mathcal{L}$-structures in an evident way. 
The universe of a graph $G$ is $V(G)$, and for each 
$1\leq i \leq c$ and any $x_1,\dots, x_{a_i}$ we say 
\[R_i(x_1,\dots,x_{a_i}) \text{ if } [x_1,\dots, x_{a_i}]\in H_i(G).\] 
That is, variables in $L$ are interpreted as vertices in $G$ and
relations are interpreted as edges. By definition, all graphs $G$ 
satisfy the formulas
\begin{itemize}
	\item Symmetry formulas:
	\[ S_g: =\big( R_i(x_1\dots,x_{a_i})\iff R_i(x_{j_1}\dots,x_{j_{a_i}})\big) ,\]
	where the the index $j_k$ is $g(k)$, and $g$ is an element from $\Phi_i$.
	\item Anti-reflexivity formulas:
	\[A_{i,(j,l)}:=\big(R_i(x_1\dots,x_{a_i})\implies 
	\neg(x_j= x_l)\big),\]
	where $(j,l)\in A_i$.
\end{itemize}

@@

From now on we will adopt the following two conventions:
\begin{itemize}
	\item We will always work in the random model $HG(n,p(n))$, where
	$p(n)=p(n,\beta)$ is defined as the tuple
	$(\beta_1 n^{1-a_1},\dots, \beta_c n^{1-a_c})$, and
	$\beta=(\beta_1,\dots, \beta_c)$.
	The symbols $\beta_1,\dots, \beta_c$
	denote positive real variables,
	but for the most part we will 
	treat them as arbitrary positive real constants.
	\item The first order language we will always refer to is the one 
	defined in last section, $\mathcal{L}$. 
\end{itemize}

@@

We define $\Lambda$ and $M$ as the minimal families
of expressions with arguments $\beta_1,\dots, \beta_c$ that satisfy the conditions:
\begin{itemize}
	\item $1\in \Lambda$
	\item For any $\mu\in M$ and any $n\in \N$ both
	$\mathrm{Poiss}_{\mu}(n)$ and $\mathrm{Poiss}_\mu(\geq n)$ are in $\Lambda$
	\item For any $\lambda_1,\lambda_2 \in \Lambda$, the
	product $\lambda_1\lambda_2$ belongs to $\Lambda$ as well.
	\item For any $l,a,i\in \N$ with $1\leq i \leq c$,
	$a > 0$, and $\lambda_1,\dots, \lambda_l\in \Lambda$,
	the expression $\frac{\beta_i \prod_{j=1}^{l}\lambda_j}{a}$
	belongs to $M$.		
\end{itemize}
We also define another families $\widehat{\Theta}$ and $\Theta$ of expressions with arguments
$\beta_1,\dots, \beta_c$ as the minimal one satisfying:
\begin{itemize}
	\item For any $l,s,a\in \N$ with $a>0$ and any 
	non-necessarily-different $\lambda_1,\dots, \lambda_l\in \Lambda$,
	$\alpha_1,\dots,\alpha_s\in \{\beta_1,\dots \beta_c\}$,
	the expression 
	\[\frac{\left(\prod_{i=1}^{l} \lambda_i \right)
		\left( \prod_{i=1}^{s}\alpha_i \right)}{a}\] 
	lies in $\widehat{\Theta}$.
	\item For any $g\in \widehat{\Theta}$ and any $n\in \N$ both
	$\mathrm{Poiss}_g(n)$ and $\mathrm{Poiss}_g(\geq n)$ are in $\Theta$.
	\item For any $\theta_1, \theta_2\in \Theta$ the product
	$\theta_1\theta_2$ belongs to $\Theta$ as well.
\end{itemize}

For any first order sentence $\phi$
we will use the notation 
$\mathrm{Pr}_n(\phi):=\mathrm{Pr}(HG(n,p(n))\models \phi)$. \par

@@

\begin{theorem}\label{thm:THEONE} 
	Let $\beta=(\beta_1,\dots, \beta_c)$, and let $\psi$ be a 
	F.O sentence in $\mathcal{L}$. Then the function
	\[
	\mathfrak{F}(\beta):=\Ln \mathrm{Pr}(\, HG(n,p(\beta,n))\, \models \, \psi \,)
	\]
	is well defined for all values of $\beta$ and it is a finite sum 
	of expressions in $\Theta$.
\end{theorem}

@@

We define a distance over arbitrary graphs $G$. For each $x, y\in V(G)$,
\[ d(x,y)= \min_{\substack{H \text{ subgraph of } G\\ 
H \text{ connected }\\
x,y\in V(H)}} |V(H)| - 1 .\]
That is, the distance between $x$ and $y$ is the minimum size
of a connected graph $H$ containing both. 
If such graph does not exist we define $d(x,y)=\infty$.
This definition extends naturally to subsets $X,Y\subseteq V(G)$:
\[ d(X,Y)=\min_{\substack{x\in X\\ y\in Y}} d(x,y).\]
As usual, when $X=\{x\}$ we will omit the brackets and write
$d(x,Y)$ instead of $d(\{x\},Y)$, for example.  \par

\begin{definition} 
	A path between $x$ and $y$ is a connected graph containing both 
	$x$ and $y$	that is minimal among the ones with those properties. 
\end{definition}

\begin{proposition}\label{prop:pathform}
	A path $P$ between $x$ and $y$ in a graph $G$ is a union of 
	edges $e_1,\dots, e_l\in H(G)$ such that
	\begin{itemize}
		\item $x$ only belongs to
		$e_1$ and $y$ only belongs to $e_l$.
		\item For any $1\leq j < i\leq l$, $e_i$ intersects
		$e_j$ if and only if $i=j+1$.
	\end{itemize}  	
\end{proposition}
\begin{proof}[Sketch of the proof.]
	Proceed by induction on the number of edges.
\end{proof}

\begin{definition}
	The likelihood $L(G)$ of an hypergraph $G=(V,E_1,\dots,E_l)$ is
	the number
	\[ \left( |V(G)| - \sum_{i=1}^l |H_i(G)|(a_i-1)\right).\]
	An hypergraph is $L$-balanced if it contains no subgraph with
	less likelihood than itself. 
\end{definition}

@@

\begin{proposition} \label{prop:balanced}
	Any connected graph $G$ is $L$-balanced. 
\end{proposition}
\begin{proof}
	\begin{proof}[Sketch of the proof.]
		Suppose $G$ is non-empty.
		The proof is by induction on the number of edges in $G$. 
		\item If $G$ has zero edges it is an isolated vertex so the statement is true.
		\item Suppose that $G$ has $m>0$ edges. Choose a vertex $x\in V(G)$ and
		an edge $e\in H(G)$ such that the distance from the $x$ to $e$ is maximum. 
		The subgraph $F$ obtained from cutting $e$ must be connected, so by the
		induction hypothesis $F$ is balanced. 
		The original graph $G$ was connected, 
		so $e$ must intersect $F$ in at least one vertex, and
		\begin{equation}\label{eqn:balanced}
		L(G)\leq L(F).
		\end{equation}
		Suppose that $G$ contains a sub-graph $G_2$ with 
		$L(G_2)<L(G)$. Then, as $F$ is $L$-balanced and
		\ref{eqn:balanced} holds, $e\in H(G_2)$. Call $F_2$
		to the result of cutting $e$ in $G_2$. Then
		$F_2$ is a subgraph of $F$ and one can check
		\[ L(F_2) - L_(G_2) \leq L(F)-L(G).\]
		In consequence, as $L(F)-L(F_2)$ is non-positive, 
		so is $L(G)-L(G_2)$, arriving at a contradiction.
	\end{proof}
\end{proof}

@@

\begin{proposition}
	A cycle $G$ is either:
	\begin{itemize}
		\item[(1)] An edge $e$ where exactly one vertex appears exactly
		twice.
		\item[(2)] A path $P$, with $L(P)=1$
		between two vertices $x$ and $y$, with
		$L(P)=1$, together with an edge $e$ that intersects $P$ exactly
		in $x$ and $y$.
	\end{itemize}
\end{proposition} 
\begin{proof}[Sketch of the proof.]
	\item If $G$ contains only one edge then (1) holds necessarily.

@@

\begin{remark}
	The $k$-type of an edge $e$ in a rooted tree $T_1$
	does not only depend on the edge $e$. It also contains some 
	non-local information about $T_1$. Namely, it contains the 
	$\morph{k}$ classes of the trees ``hanging'' from $e$.
	If $T_2$ is a sub-tree of $T_1$ also containing $e$, the $k$-type
	of $e$ in $T_2$ may differ from the $k$-type of $e$ in $T_1$.
	However we will usually refer to ``the $k$-type of $e$'' instead of
	``the $k$-type of $e$ in $T_1$'' when the ``ambient tree'' is clear
	from context.   
\end{remark}

@@

\begin{lemma} \label{lem:lessktree}
	Let $(T_1,x)$ and $(T_2,y)$ be rooted trees such that for some $k>0$, 
	$T_1 \morph{k} T_2$. Then, for any $0\leq j < k$, it is satisfied
	$T_1 \morph{j} T_2$.
\end{lemma}
\begin{proof}
	Fix $j<k$. The proof is by induction on the radius of $T_1$ and $T_2$.
	\begin{itemize}[leftmargin=*]
		\item If they have radius $0$ they are $j$-morphic by definition. 
		\item Suppose now that they have radius $r$. Let $e_1$ be an initial edge 
		in $T_1$ and $e_2$ be an initial edge in $T_2$ such that $e_1 \morph{k} e_2$.
		Then, because of the definition of $k$-type,
		$e_1$ and $e_2$ have the same color $l$ and we can write
		$e_1=[x_1,\dots,x_{c_l}]$,
		$e_2=[y_1,\dots,y_{c_l}]$ for some vertices such that for all
		$i=1,\dots, a_l$, either 
		\begin{itemize}
			\item $Tree(x_i)\morph{k} Tree(y_i)$, and because of the induction hypothesis $Tree(x_i)\morph{j} Tree(y_i)$. 
			\item Or both $x_i=x$ and $y_i=y$ are the roots of $T_1$ and $T_2$.
		\end{itemize}
		In consequence $e_1\morph{j} e_2$. Then, the $\morph{k}$ relation
		implies the $\morph{j}$ relation on edges with radius less than $r$.
		Thus, for any $j$-type of edges $E$ there are 
		$k$-types of edges $E_1, \dots, E_m$ such that for all rooted trees $T$
		\[ \langle T,E\rangle = \sum_{i=1}^{m} \langle T,E_i \rangle .\]
		In particular
		\[\langle T_g , E\rangle =\sum_{i=1}^m \langle T_g, E_i \rangle 
		\quad \text{, for } 	g=1,2.\]	
	 	Both of the above sums coincide if all the numbers
	 	$\langle T_1, E_i\rangle$'s are not greater than $k$ or
	 	both sums are greater than $k$ otherwise. As $j\leq k$ this implies
	 	\[\text{Either }\langle T_1, E \rangle =\langle T_2,E \rangle \text{ or } \langle T_1,E \rangle ,\langle T_2,C \rangle \geq j+1 ,\]
	 	and $T_1\morph{j}T_2$. 		
	\end{itemize}
\end{proof}

@@

\begin{definition} 
	Two graphs with constants $G_1$ and $G_2$ are $k$-agreeable if for each $k$-morphism 
	equivalence class $C$ they either have the same number of connected components of type
	$C$ or they both have no less than $k+1$ components of type $C$. 
\end{definition}

@@

For reference about E.F. games see \cite{finitemodeltheory1} and \cite{finitemodeltheory2}.

>>>>

\usepackage[left=3.7cm,right=3.7cm,top=3cm,bottom=3cm]{geometry}
\usepackage{graphicx}
%%BeginIpePreamble
\usepackage{amssymb,mathtools, amsmath, amsfonts, amsthm}
%%EndIpePreamble
%\usepackage{color}
\usepackage{float}
\usepackage{hyperref}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{chngcntr}
\usepackage{cleveref}
\usepackage{pdfpages}

@@

The original goal was to give a generalization for the case of uniform hypergraphs
but the techniques used for that case also allowed for a generalization for 
hypergraphs with multiple edge sets of different sizes.

@@

This way, asymptotically, trees of any kind occur arbitrarily often in a typical
random graph. Given the locality of first order logic up to a fixed 
quantifier rank this is tells us that the only way 
to distinguish large random graphs in $G(n,\beta/n)$ under first order logic
(up to a fixed quantifier rank) 
is by looking at the neighborhoods of their small cycles. \par

It is given a finite and exhaustive classification of the uni-cyclic
graphs with less than a given radius. After that it is shown that
if two graphs $G_1$ and $G_2$ contain the same number 
of uni-cycles of each type or they
contain a large enough number of those then they are a.a.s equivalent
under first order logic up to some quantifier rank $k$. This is done by giving
a winning strategy for duplicator in the Ehrenfeucht Fraisse game of $k$ rounds played on 
$G_1$ and $G_2$.\par

Finally it is shown that the quantities of uni-cycles of each type in a random graph
are asymptotically distributed like independent Poisson variables 
whose means are given by some other nested ``Poisson expressions" that correspond 
to the trees hanging from each uni-cycle.

@@

Given a natural number $n$, we will use the notation
$[n]:= \{1,\dots, n\}$. We will denote by $S_n$
the symmetric group on $[n]$, and by $\Delta_n$ the 
diagonal set $\{(a,a)\in [n]^2 \}$. \par
Given a set $X$, then $S_n$ acts on $X^n$ in a natural way. That is, 
given $g\in S_n$ and $(x_1, \dots, x_n)$ one can define
\[ g \cdot (x_1,\dots,x_n) =(y_1,\dots,y_n), \]
where $y_{g(i)}=x_i$ for all $1\leq i\leq n$.\par
Given $\Phi$ a subgroup of $S_n$ we will denote by $X^n/\Phi$
the orbit set associated to the action of $\Phi$ over $X^n$.\par
We will use the notation $[x_1,\dots,x_n]$ to refer to the 
equivalence class of the $n$-tuple $(x_1,\dots, x_n)$ in 
any sort of quotient $X^n/\Phi$. That is, while the notation
$(x_1,\dots, x_n)$ will be reserved to ordered $n$-tuples, 
$[x_1,\dots,x_n]$ will denote an ordered $n$-tuple modulo the
action of some arbitrary group of permutations. Which group is this 
will depend solely on the ambient set where $[x_1,\dots,x_n]$ is
considered.
\par

\begin{definition}
	Let $n,a\in \N$, with $a\geq 2$, let $\Phi$ be a subgroup of $S_a$, and let
	$A$ be a subset 
	\[ A\subseteq [a]^2 \setminus \Delta_a.\]
	The total edge set $\mathcal{H}_{(a,\Phi, A)}(n)$ of size $a$, 
	symmetry group $\Phi$ and
	restrictions $A$ on $n$ elements is the set:
	\[  \mathcal{H}_{(a,\Phi, A)}(n)= ([n]^a/\Phi) \, \,
	\setminus \{\,  [x_1, \dots,x_a] \in [n]^a/\Phi  \, \, 
	| \, \, x_i=x_j \, \text{for some } (i,j)\in A \} \]
\end{definition}

@@

\begin{definition} 
	Let $p=(p_1,\dots, p_c)$, where all $p_i$'s are real numbers
	between $0$ and $1$. 
	The random model $HG(n,p)$ with edge colors $1,\dots,c$,
	sizes $a_1,\dots,a_c$, 
	symmetry groups $\Phi_1,\dots,\Phi_c$ 
	and restrictions $A_1,\dots,A_c$, is the one that
	assigns to each graph $G=([n], H_1,\dots, H_c)$
	probability
	\[ \mathrm{Pr}(G)=\prod_{i=1}^{c} p_i^{|H_i|}(1-p_i)^{|\mathcal{H}_{(a_i,\Phi_i,A_i)}(n)|-|H_i|}.	
	\]
	Equivalently, this is the probability space obtained by 
	assigning to each colored edge 
	$e\in \mathcal{H}_{(a_i,\Phi_i,A_i)}(n)$ probability $p_i$ independently. 
\end{definition}

For the rest of the work we will consider  fixed
\begin{itemize}
	\item the total number of colors $c$,	
	\item the sizes $a_1,\dots,a_c$,
	\item the symmetry group $\Phi_1,\dots, \Phi_c$ and,
	\item the restrictions $A_1,\dots,A_c$.
\end{itemize}
When we say ``graph'' from now on what
we will mean is ``hiper-graph with edge colors 
$1,\dots, c$, sizes $a_1,\dots,a_c$, 
symmetry groups $\Phi_1,\dots,\Phi_c$ and 
restrictions $A_1,\dots,A_c$''. \par

Given a graph $G=([n], H_1,\dots, H_c)$ we will denote by
$H_i(G)$ the edge set $H_i$, and by $V(G)$ the vertex set
$[n]$. Also, we will write $H(G)$ to denote the 
disjoint union of colored sets $\cup_{i=1}^c H_i$.
This way, an edge $e\in H(G)$ with color $i$ is an element
$[x_1,\dots,x_{a_i}]\in H_i(G)$, and the $x_i$'s are vertices
belonging to $V(G)$. \par
Given a set of vertices, $X\subseteq V(G)$, we will denote
the by $G[X]$ the induced sub-graph on $X$. \par
As usual, we will sometimes treat edges as sets of vertices
rather than ``tuples modulo the action of some permutation group''. 
This way, expressions like $e_1\cap e_2$ for $e_1, e_2\in H(G)$
will make sense and mean 
``the set of vertices that occupy some place in $e_1$ 
and in $e_2$".\par
Some other times we will treat edges $e\in H(G)$
as sub-graphs of $G$ in the natural way. That is, the subgraph
denoted by $e$ is the one whose vertex set is $e$-i.e., the vertices in $e$-
and whose only edge is $e$. 
This way, when we have some edges $e_1,\dots, e_l\in H(G)$ 
it will make sense to talk
about the subgraph $\cup_{i=1}^l e_i$, which is the graph
whose vertex set is the set of vertices belonging to any of the $e_i$'s, 
and whose edges are exactly the $e_i$'s. In spite of these
abuses of notation the ``type'' of any ``term'' involving edges 
should be derivable from the context. \par
Another usual abuse of notation we will make is to sometimes
treat graphs as their underlying vertex sets. Hence,
expressions defined for sets of vertices will also be defined 
for graphs.

@@

Graphs are $\mathcal{L}$-structures in an evident way. 
The universe of a graph $G$ is $V(G)$, and for each 
$1\leq i \leq c$ and any $x_1,\dots, x_{a_i}$ we say 
\[R_i(x_1,\dots,x_{a_i}) \text{ if } [x_1,\dots, x_{a_i}]\in H_i(G).\] 
That is, variables in $L$ are interpreted as vertices in $G$ and
relations are interpreted as edges. By definition, all graphs $G$ 
satisfy the formulas
\begin{itemize}
	\item Symmetry formulas:
	\[ S_g: =\big( R_i(x_1\dots,x_{a_i})\iff R_i(x_{g(1)}\dots,x_{g(a_i)})\big) ,\]
	where $g$ is an element from $\Phi_i$.
	\item Anti-reflexivity formulas:
	\[AR_{i,(j,l)}:=\big(R_i(x_1\dots,x_{a_i})\implies 
	\neg(x_j= x_l)\big),\]
	where $(j,l)\in A_i$.
\end{itemize}

@@

From now on we will adopt the following two conventions:
\begin{itemize}
	\item We will always work in the random model $HG(n,p(n))$, where
	$p(n)=p(n,\beta)$ is defined as the tuple
	$(\beta_1 n^{1-a_1},\dots, \beta_c n^{1-a_c})$, and
	$\beta=(\beta_1,\dots, \beta_c)$.
	The symbols $\beta_1,\dots, \beta_c$
	denote non-negative real variables,
	but for the most part we will 
	treat them as non-negative positive real constants.
	\item The first order language we will always refer to is the one 
	defined in last section, $\mathcal{L}$. 
\end{itemize}

@@

We define $\Lambda$ and $M$ as the minimal families
of expressions with arguments $\beta_1,\dots, \beta_c$ that satisfy the conditions:
\begin{itemize}
	\item $1\in \Lambda$
	\item For any $\mu\in M$ and any $n\in \N$ both
	$\mathrm{Poiss}_{\mu}(n)$ and $\mathrm{Poiss}_\mu(\geq n)$ are in $\Lambda$
	\item For any $\lambda_1,\lambda_2 \in \Lambda$, the
	product $\lambda_1\lambda_2$ belongs to $\Lambda$ as well.
	\item For any $b,i\in \N$ with $1\leq i \leq c$,
	$b > 0$, and $\lambda_1,\dots, \lambda_{a_i-1}\in \Lambda$,
	the expression $\frac{\beta_i \prod_{j=1}^{a_i-1}\lambda_j}{b}$
	belongs to $M$.		
\end{itemize}
We also define another families $\widehat{\Theta}$ and $\Theta$ of expressions with arguments
$\beta_1,\dots, \beta_c$ as the minimal ones satisfying:
\begin{itemize}
	\item For any $l,s,b\in \N$ with $b>0$ and any 
	non-necessarily-different $\lambda_1,\dots, \lambda_l\in \Lambda$,
	$\alpha_1,\dots,\alpha_s\in \{\beta_1,\dots \beta_c\}$,
	the expression 
	\[\frac{\left(\prod_{i=1}^{l} \lambda_i \right)
		\left( \prod_{i=1}^{s}\alpha_i \right)}{b}\] 
	lies in $\widehat{\Theta}$.
	\item For any $u\in \widehat{\Theta}$ and any $n\in \N$ both
	$\mathrm{Poiss}_u(n)$ and $\mathrm{Poiss}_u(\geq n)$ are in $\Theta$.
	\item For any $\theta_1, \theta_2\in \Theta$ the product
	$\theta_1\theta_2$ belongs to $\Theta$ as well.
\end{itemize}
\vspace{1 em}

For any first order sentence $\varphi$
we will use the notation 
\[\mathrm{Pr}_n(\varphi):=\mathrm{Pr}(HG(n,p(n))\models \varphi).\]
\par

@@

\begin{theorem}\label{thm:THEONE} 
	Let $\beta=(\beta_1,\dots, \beta_c)$, and let $\varphi$ be a 
	F.O sentence in $\mathcal{L}$. Then the function
	\[
	\mathfrak{F}(\beta):=\Ln \mathrm{Pr}(\, HG(n,p(\beta,n))\, \models \, \varphi \,)
	\]
	is well defined for all values of $\beta$ and it is a finite sum 
	of expressions in $\Theta$.
\end{theorem}

@@

We define a distance over arbitrary graphs $G$. For each $x, y\in V(G)$,
\[ d(x,y)= \min_{\substack{H \text{ subgraph of } G\\ 
H \text{ connected }\\
x,y\in V(H)}} |V(H)| - 1 .\]
That is, the distance between $x$ and $y$ is the minimum
of a connected graph $H$ containing both, minus one. 
If such graph does not exist we define $d(x,y)=\infty$.
This definition extends naturally to subsets $X,Y\subseteq V(G)$:
\[ d(X,Y)=\min_{\substack{x\in X\\ y\in Y}} d(x,y).\]
As usual, when $X=\{x\}$ we will omit the brackets and write
$d(x,Y)$ instead of $d(\{x\},Y)$, for example.  \par

\begin{definition} 
	A path between $x$ and $y$ is minimal graph among the connected graphs
	containing both $x$ and $y$. 
\end{definition}

\begin{proposition}\label{prop:pathform}
	A path $P$ between $x$ and $y$ in a graph $G$ is a union of 
	edges $e_1,\dots, e_l\in H(G)$ such that
	\begin{itemize}
		\item $x$ only belongs to
		$e_1$ and $y$ only belongs to $e_l$.
		\item For any $1\leq j < i\leq l$, $e_i$ intersects
		$e_j$ if and only if $i=j+1$.
	\end{itemize}  	
\end{proposition}
\begin{proof}[Sketch of the proof.]
	Order the edges of the path in a way that (1) $x$ belongs to the first and $y$ to the last, and
	(2) each edge intersects the previous one and the next one. Such ordering exists because a path is connected.
	If an edge intersects any other edge than its previous one or next one then we can remove some edge of the path,
	contradicting the minimality condition.
\end{proof}

\begin{definition}
	The likelihood $L(G)$ of an hypergraph $G=(V,H_1,\dots,H_c)$ is
	the number
	\[ \left( |V(G)| - \sum_{i=1}^c |H_i|(a_i-1)\right).\]
	An hypergraph is $L$-balanced if it contains no subgraph with
	less likelihood than itself. 
\end{definition}

@@

\begin{proposition} \label{prop:balanced}
	Any connected graph $G$ is $L$-balanced. 
\end{proposition}
	\begin{proof}
		Suppose $G$ is non-empty.
		The proof is by induction on the number of edges in $G$. 
		\item If $G$ has zero edges it is an isolated vertex so the statement is true.
		\item Suppose that $G$ has $m>0$ edges. Choose a vertex $x\in V(G)$ and
		an edge $e\in H(G)$ such that the distance from the $x$ to $e$ is maximum. 
		The subgraph $F$ obtained from cutting $e$ must be connected, so by the
		induction hypothesis $F$ is L-balanced. 
		The original graph $G$ was connected, 
		so $e$ must intersect $F$ in at least one vertex, and
		\begin{equation}\label{eqn:balanced}
		L(G)\leq L(F).
		\end{equation}
		Suppose that $G$ contains a sub-graph $G_2$ with 
		$L(G_2)<L(G)$. Then, as $F$ is $L$-balanced and
		\ref{eqn:balanced} holds, $e\in H(G_2)$. Call $F_2$
		to the result of cutting $e$ in $G_2$. Then
		$F_2$ is a subgraph of $F$ and one can check
		\[ L(F_2) - L(G_2) \leq L(F)-L(G).\]
		In consequence, as $L(F)-L(F_2)$ is non-positive, 
		so is $L(G)-L(G_2)$, arriving at a contradiction.
	\end{proof}

@@

\begin{proposition}
	A cycle $G$ is either:
	\begin{itemize}
		\item[(1)] An edge $e$ where exactly one vertex appears exactly
		twice.
		\item[(2)] A path $P$, with $L(P)=1$
		between two vertices $x$ and $y$, 
		together with an edge $e$ that intersects $P$ exactly
		in $x$ and $y$.
	\end{itemize}
\end{proposition} 
\begin{proof}[Sketch of the proof.]
	\item If $G$ contains only one edge then (1) holds necessarily.

@@

\begin{remark}
	The $k$-type of an edge $e$ in a rooted tree $T_1$
	does not only depend on the edge $e$. It also contains some 
	non-local information about $T_1$. Namely, it contains the 
	$\morph{k}$ classes of the trees ``hanging'' from $e$.
	If $T_2$ is a sub-tree of $T_1$ also containing $e$, the $k$-type
	of $e$ in $T_2$ may differ from the $k$-type of $e$ in $T_1$.
	However we will usually refer to ``the $k$-type of $e$'' instead of
	``the $k$-type of $e$ in $T_1$'' when the ``ambient tree'' is clear
	from the context.   
\end{remark}

@@

\begin{lemma} \label{lem:lessktree}
	Let $(T_1,x)$ and $(T_2,y)$ be rooted trees such that for some $k>0$, 
	$T_1 \morph{k} T_2$. Then, for any $0\leq j < k$, it is satisfied
	$T_1 \morph{j} T_2$.
\end{lemma}
\begin{proof}
	Fix $j<k$. The proof is by induction on the radius of $T_1$ and $T_2$.
	\begin{itemize}[leftmargin=*]
		\item If they have radius $0$ they are $j$-morphic by definition. 
		\item Suppose now that they have radius $r$. Let $e_1$ be an initial edge 
		in $T_1$ and $e_2$ be an initial edge in $T_2$ such that $e_1 \morph{k} e_2$.
		Then, because of the definition of $k$-type,
		$e_1$ and $e_2$ have the same color $l$ and we can write
		$e_1=[x_1,\dots,x_{a_l}]$,
		$e_2=[y_1,\dots,y_{a_l}]$ for some vertices such that for all
		$i=1,\dots, a_l$, either 
		\begin{itemize}
			\item $Tree(x_i)\morph{k} Tree(y_i)$, and because of the induction hypothesis $Tree(x_i)\morph{j} Tree(y_i)$. 
			\item Or both $x_i=x$ and $y_i=y$ are the roots of $T_1$ and $T_2$.
		\end{itemize}
		In consequence $e_1\morph{j} e_2$. Then, the $\morph{k}$ relation
		implies the $\morph{j}$ relation on edges with radius less than $r$.
		Thus, for any $j$-type of edges $E$ there are 
		$k$-types of edges $E_1, \dots, E_m$ such that for all rooted trees $T$ of radius $r$
		\[ \langle T,E\rangle = \sum_{i=1}^{m} \langle T,E_i \rangle .\]
		In particular
		\[\langle T_g , E\rangle =\sum_{i=1}^m \langle T_g, E_i \rangle 
		\quad \text{, for } 	g=1,2.\]	
	 	Both of the above sums coincide if all the numbers
	 	$\langle T_1, E_i\rangle$'s are not greater than $k$ or
	 	both sums are greater than $k$ otherwise. As $j\leq k$ this implies
	 	\[\text{Either }\langle T_1, E \rangle =\langle T_2,E \rangle \text{ or } \langle T_1,E \rangle ,\langle T_2,C \rangle \geq j+1 ,\]
	 	and $T_1\morph{j}T_2$. 		
	\end{itemize}
\end{proof}

@@

@@

For reference about Ehrenfeucht Fraisse games see \cite{finitemodeltheory1} and \cite{finitemodeltheory2}.
